Follow-up Samples

### 2017-07-21 submission/manuscript.tex

\textbf{Results.} There were small changes to the bacterial community
associated with adenoma or advanced adenoma groups and large changes
associated with the carcinoma group. The communities from patients with
carcinomas changed significantly more than those with adenoma following
treatment (P-value \textless{} 0.001). There was no significant change
in the microbiota between patients with adenoma and advanced adenoma, or
between patients with advanced adenoma and carcinoma (P-value
\textgreater{} 0.05). Although treatment was associated with
intrapersonal changes, the change in the abundance of individual OTUs to
treatment was not consistent within diagnosis groups (P-value
\textgreater{} 0.05). Because the distribution of OTUs across patients
and diagnosis groups was irregular, we used the Random Forest machine
learning algorithm to try and identify groups of OTUs that would allow
us to successfully distinguish between pre and post-treatment samples
for each of the diagnosis groups. Although the three models could
successfully differentiate between the pre and post-treatment samples,
there was little overlap between the OTUs that were indicative of
treatment and randomization of the diagnosis labels recapitulated the
actual models AUCs. Next, we used a larger cohort that contained
individuals with normal colons and those with adenomas, advanced
adenomas, and carcinomas to determine whether individuals who underwent
treatment were more likely to have OTUs associated with normal colons.
We again built Random Forest models and measured the change in the
positive probability of having one of the three diagnoses to assess
whether these changes were still similar to their original diagnosis or
not. Only patients who had carcinomas had a change that was towards a
microbial milieu that resembles the normal colon (P-value \textless{}
0.05). Finally, we tested whether the type of treatment impacted the
microbiota of those diagnosed with carcinomas and were unable to detect
any significant differences in characteristics of these communities
between individuals treated with surgery alone and those treated with
chemotherapy or chemotherapy and radiation (P-value \textgreater{}
0.05).

>>>>

\textbf{Results.} There were small changes to the bacterial community
associated with adenoma or advanced adenoma and large changes associated
with the carcinoma. The communities from patients with carcinomas
changed significantly more than those with adenoma following treatment
(P-value \textless{} 0.001). Although treatment was associated with
intrapersonal changes, the change in the abundance of individual OTUs to
treatment was not consistent within diagnosis groups (P-value
\textgreater{} 0.05). Because the distribution of OTUs across patients
and diagnosis groups was irregular, we used the Random Forest machine
learning algorithm to try and identify groups of OTUs that would allow
us to distinguish between pre and post-treatment samples for each of the
diagnosis groups. Although the three models could differentiate between
the pre and post-treatment samples, there was little overlap between the
OTUs that were indicative of treatment. Next, we used a larger cohort
that contained individuals with normal colons and those with adenomas,
advanced adenomas, and carcinomas to determine whether individuals who
underwent treatment were more likely to have OTUs associated with normal
colons. We again built Random Forest models and measured the change in
the positive probability of having one of the three diagnoses to assess
whether these changes were similar to their original diagnosis or not.
Patients who had carcinomas changed towards a microbial milieu that
resembles the normal colon (P-value \textless{} 0.05). Finally, we were
unable to detect any significant differences in the microbiota between
individuals treated with surgery alone and those treated with
chemotherapy or chemotherapy and radiation (P-value \textgreater{}
0.05).
